% =========================
% Task 1
% =========================

\section{Task 1: Transfer Learning for Pet Classification}

In this task, we study image classification on the Oxford-IIIT Pet Dataset using deep neural networks. The primary objective is to investigate the effectiveness of transfer learning by fine-tuning an ImageNet-pretrained ResNet-34 model. The original classification head is replaced to match the target category number, while the pretrained backbone is fine-tuned with a smaller learning rate.

Based on the baseline model, we further analyze the influence of several important hyperparameters, including optimizer type, backbone learning rate, classifier learning rate, batch size, and weight decay. An ablation study is also conducted to compare pretrained initialization with random initialization, in order to evaluate the contribution of pretrained representations.

In addition, attention mechanisms are incorporated into the baseline architecture for comparison. Specifically, SE-blocks are integrated into the ResNet model, and a lightweight Vision Transformer model is also implemented as an alternative architecture. Experimental results are evaluated primarily using classification accuracy on the validation and test sets.


\input{sections/task1/Dataset and Data Preprocessing.tex}
\input{sections/task1/Baseline Transfer Learning Model.tex}
\input{sections/task1/Attention-Based Architectures.tex}
\input{sections/task1/Training Pipeline.tex}
\input{sections/task1/Hyperparameter Analysis.tex}
\input{sections/task1/Pretraining Ablation Study.tex}

% =========================
% Experimental Results
% =========================

\subsection{Experimental Results}

% TODO:
% Main quantitative results table
% Accuracy comparison
% Baseline vs SE vs ViT-Tiny
% Pretrained vs Random Initialization

% =========================
% Visualization
% =========================

\subsection{Training Visualization}

% TODO:
% Loss curves
% Accuracy curves
% Hyperparameter search visualization
% Ablation visualization

% =========================
% Discussion
% =========================

\subsection{Discussion}

% TODO:
% Observations about transfer learning
% Effect of SE block
% Hyperparameter sensitivity
% Why pretrained initialization helps
% Discussion about initialization strategy

% =========================
% Summary
% =========================

\subsection{Summary}

% TODO:
% Summarize key findings of Task 1
